---
title: "ECON 432 Homework 1"
author: "Madelyn Caufield"
date: "`r format(Sys.Date(),'%b %d,%Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    number_sections: yes
    theme: united
    highlight: tango
---

```{r}
getwd()
setwd("/Users/madelyncaufield/Desktop/ECON432 Data Science for Financial Engineering/Homework 1")
getwd()
```

```{r}
df <- read.csv("sbuxPrices.csv")
head(df)
```

```{r}
library(dplyr)

sbux.df = df %>%
          select(Date, Adj.Close)
```

# Part II
## Question 1
```{r}
head(sbux.df)
```
This gives us the first 6 rows of the data.

```{r}
tail(sbux.df)
```
This gives us the last 6 rows of the data.

```{r}
sbuxPrices.df = sbux.df[, "Adj.Close", drop=FALSE]
sbuxPrices.df
```
This gives us just the column for adjusted closing prices and drops the column Date column to make a new data frame called sbuxPrices.df. 

```{r}
rownames(sbuxPrices.df) = sbux.df$Date
rownames(sbuxPrices.df)
```
This converted the Date column to row names for the sbuxPrices.df data frame. 

```{r}
head(sbuxPrices.df)
```
This gave us the first 6 rows of the adjusted closing price column with the new row names (dates from above) from the sbuxPrices.df data frame.

```{r}
sbux.df[101:132,]
```
This gave us only rows 101-132 for the sbux.df data frame. 

## Question 2
```{r}
class(sbux.df$Date)
sbux.df$Date <- as.Date(sbux.df$Date,format = '%Y-%m-%d')
class(sbux.df$Date)
```

```{r}
plot(Adj.Close~Date, sbux.df, type='l',col="blue",xlab = 'Date',ylab='Starbucks Prices',
     main='Monthly Prices of Starbucks Shares')
T <- length(sbux.df$Date)
ticks <- seq(sbux.df$Date[1], sbux.df$Date[T], by = "quarters")
axis(1, at = ticks, labels = FALSE)
legend('topleft', names(sbuxPrices.df), col= "blue", lty=1, cex=.65)
```

## Question 3
### Monthly Simple Return
```{r}
sbux.df$ret[2:T] <-  sbux.df$Adj.Close[2:T]/sbux.df$Adj.Close[1:(T-1)] - 1
head(sbux.df$ret)
```

```{r}
library(ggplot2)

df.g <- sbux.df[,c('Date','ret')] 
df.g <- df.g[-c(1),]

head(df.g)
```

### Monthly Simple Return Plot
```{r}
g <- ggplot(df.g, aes(x = Date, y = ret)) + 
  geom_line(color="brown1") +  
  ylab("Starbucks Simple Returns") +
  xlab("Date (03/01/1998 - 03/01/2013)") +
  ggtitle("Monthly Simple Returns of Starbucks Shares")
g
```

### Monthly Coninuously Compounded Return
```{r}
sbux.df$cc.ret[2:T] <-  log(sbux.df$Adj.Close[2:T]/sbux.df$Adj.Close[1:(T-1)])
head(sbux.df$cc.ret)

df.g$cc.ret <- sbux.df$cc.ret[2:T]
head(df.g)
```

### Monthly Continuously Compounded Return Plot
```{r}
g <- ggplot(df.g, aes(x = Date, y = cc.ret)) + 
  geom_line(color="lightseagreen") +  
  ylab("Starbucks Continuously Compounded Returns") +
  xlab("Date (03/01/1998 - 03/01/2013)") +
  ggtitle("Monthly Continuously Compounded Returns of Starbucks Shares")
g
```

### Plot of the Monthly Simple Return and Continuously Compounded Return
```{r}
library(reshape2)
d <- melt(df.g, id.vars="Date")

ggplot(d, aes(Date, value, col=variable)) + 
  geom_line() +
  ggtitle("Monthly Simple Returns and CC Returns of Starbucks Shares")
```










